---
title: "Project Report"
author: "Meghana Inamanamelluri"
date: "12/9/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include=FALSE}
#install.packages("tidyverse")
#install.packages("rio")
```


```{r include=FALSE}
library(resampledata)
library(tidyverse)
library(rio)
library(ggplot2)
```


## Do Children Copy an Expert or a Majority? Examining Selective Learning in Instrumental and Normative Contexts


This study examined whether instrumental and normative learning contexts differentially influence 4- to 7-year-old children’s social learning strategies; specifically, their dispositions to copy an expert versus a majority consensus

## Experiment 1 :

In Experiment 1, they have tested whether 4-to-7-year-olds selectively copy an instrumentally competent expert over an incompetent individual.They included 4-to-7-year-old children to examine whether there is a developmental progression in the degree of copying faithfully and whether there are differences in development according to context.


### Converting and Reading Dataset1 :


```{r}
# convert .SAV file to .csv file
data_source <- "/Users/meghanainamanamelluri/Desktop/Final Project/DataSets/S1_Dataset.SAV"
convert(data_source,"/Users/meghanainamanamelluri/Desktop/Maths/DataSets/S1_Dataset.csv")
```



```{r read_data_from_csv,}
#reading the csv file
data_source <- "/Users/meghanainamanamelluri/Desktop/Maths/DataSets/S1_Dataset.csv"
s1_data <- read_csv(data_source)
glimpse(s1_data)
```
